---
title: "Thesis Analyses (IC)"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Part 1

### Intro

Let's start by loading in our packages:
```{r packages, message = FALSE}
library(tidyverse)
library(readxl)
library(rstatix)
library(ggpubr)
library(ez)
library(emmeans)
library(car)
library(Hmisc)
library(corrplot)
library(multcomp)
library(lme4)
```

Now let's take a second to think about the data we have. Overall I have seven days on the full paradigm without injection, two full days with either saline or naltrexone injection, two days with either pure fat or pure sugar only, and a final two days with each animal's favorite diet ( measured as mean average kcal eaten per day) following either saline or naltrexone injection.

My prospectus calls for 5 days of full paradigm training and 3 days of full paradigm testing, after which each animal's favorite diet was to be calculated from the 3 test days. For some reason, what we did instead was use all 7 days to calculate favorite diet. I think that's fine, but it leaves as an open question what we're calling training and what we're calling testing.

I think for now, the best thing to do is to stick as close to my prospectus as I can. I need at least 3 days of test data, so that leaves the four previous days as training data. Let's start there.

### Data Prep

Let's read in our data:

```{r read data}
#Vector for column names and diet names
column_names <- c("rat", "zero", "twenty", "forty", "sixty", "eighty", "hundred", "h2o", "day")
diet_names <- c("zero", "twenty", "forty", "sixty", "eighty", "hundred")
rat_names <- c("IC1", "IC2", "IC3", "IC4", "IC5", "IC6", "IC7", "IC8", "IC9", "IC10", "IC11", "IC12")

datap1_grams_wide <- read_excel("IC Test Data.xlsx", sheet = 14, range = "A3:I38", col_names = column_names)
datap1_kcal_wide <- read_excel("IC Test Data.xlsx", sheet = 14, range = "K3:S38", col_names = column_names)

#Convert negative values to zeroes
datap1_grams_filtered <- datap1_grams_wide
datap1_grams_filtered[datap1_grams_filtered < 0] <- 0

datap1_kcal_filtered <- datap1_kcal_wide
datap1_kcal_filtered[datap1_kcal_filtered < 0] <- 0

#Drop water for now
datap1_grams_filtered2 <- datap1_grams_filtered %>%
  subset(select = -c(h2o))

datap1_kcal_filtered2 <- datap1_kcal_filtered %>%
  subset(select = -c(h2o))

#Pivot to long format
datap1_grams <- datap1_grams_filtered2 %>%
  pivot_longer(cols = diet_names, names_to = "diet", values_to = "grams")
datap1_grams$day <- factor(datap1_grams$day,
                           levels = c("1", "2", "3"))
datap1_grams$diet <- factor(datap1_grams$diet,
                            levels = diet_names)
datap1_grams$rat <- factor(datap1_grams$rat,
                            levels = rat_names)

datap1_kcals <- datap1_kcal_filtered2 %>%
   pivot_longer(cols = diet_names, names_to = "diet", values_to = "kcals")
datap1_kcals$day <- factor(datap1_kcals$day,
                           levels = c("1", "2", "3"))
datap1_kcals$diet <- factor(datap1_kcals$diet,
                            levels = diet_names)
datap1_kcals$rat <- factor(datap1_kcals$rat,
                            levels = rat_names)
```

### Initial Vizualizations

Now just like normally, I think the best thing to do is to look at some visualizations before I do anything else

```{r initial viz}
ggboxplot(datap1_grams,
          x = "diet", y = "grams", color = "diet")

ggboxplot(datap1_kcals,
          x = "diet", y = "kcals", color = "diet")
```
This is interesting, and very different from the ID data. Here, it looks like there was a bifurcation between twenty and eighty percent as favorite diets. I already knew this since we had to give each rat it's favorite diet later in the study, but it's interesting to see it look so pronounced here.

Next let's facet this by rat to see the breakdown for each individual:

```{r initial viz 2}
ggboxplot(datap1_grams,
          x = "diet", y = "grams", color = "diet", facet.by = "rat")

ggboxplot(datap1_kcals,
          x = "diet", y = "kcals", color = "diet", facet.by = "rat")
```

Again, very much not the same as the ID data. It looks like rats run the gamut of favorite diets. These change some between datasets (which means I'll have to justify using caloric intake to choose favorite diets; shit), but not overwhelmingly. I wish we'd given DAMGO to these animals--we may have seen something really cool.

In any case, because we're including day as a dependent variable in the final analysis, I want to get a sense of how much this picture changes by day:

```{r initial viz 3}
ggboxplot(datap1_grams,
          x = "diet", y = "grams", color = "diet", facet.by = "day")

ggboxplot(datap1_kcals,
          x = "diet", y = "kcals", color = "diet", facet.by = "day")
```

So there's a healthy amount of variability between days, but not enough to significantly change the rank orderings of each diet.

### Inferential Analysis

Now for the part we've all been waiting for, the inferential analysis:

```{r hypothesis test grams}
res.aovGrams = anova_test(
  data = datap1_grams,
  dv = grams,
  wid = rat,
  within = c(diet, day),
  type = 2
)
get_anova_table(res.aovGrams)

#trying it a different way
exp2Grams <- aov(grams ~ diet*day + Error(rat/(diet*day)), data = datap1_grams)
summary(exp2Grams) 

#And a third way because the first two don't match
grams.aov <- ezANOVA(data = datap1_grams,
                    dv = .(grams),
                    wid = .(rat),
                    within = .(diet, day),
                    return_aov = TRUE,
                    type = 2)
grams.aov$ANOVA
```

The specific numbers between the alternate methods I'm using vary, but they all tell the same basic story. We have a main effect of diet, no effect of day, and no interaction. That's good, because a main effect of diet is all we were looking for. 

Let's look further into this by using pairwise within subjects t-tests to see how diet groups differed from each other:

```{r pairwise tests}
pairwiseGrams <- datap1_grams %>%
  pairwise_t_test(grams~diet, paired = TRUE,
                  p.adjust.method = "bonferroni")
pairwiseGrams
```
So from this it looks like even with Bonferroni corrections we have a decent number of significant comparisons. 80% and 20% don't actually significantly differ from each other (which is fascinating) but they're more consumed than pretty much everything else. 

And finally, let's do an explicit contrast comparing the individual diets with the blends:

```{r contrast grams 1}
#grams.em1 <- emmeans(grams.aov$aov, ~diet * day)

#contrast_names <- list(c(2, -1, -1, -1, -1, 2,
 #                        2, -1, -1, -1, -1, 2,
  #                       2, -1, -1, -1, -1, 2))
#contrast(grams.em1,
 #        method = contrast_names)
```

For completeness sake, I want to do this all over again with the calorie data

```{r hypothesis test kcals}
#trying it a different way
exp2Kcals <- aov(kcals ~ diet*day + Error(rat/(diet*day)), data = datap1_kcals)
summary(exp2Kcals) 

#And a third way because the first two don't match
kcals.aov <- ezANOVA(data = datap1_kcals,
                    dv = .(kcals),
                    wid = .(rat),
                    within = .(diet, day),
                    return_aov = TRUE,
                    type = 2)
kcals.aov$ANOVA
```

```{r pairwise tests kcals}
pairwiseKcals <- datap1_kcals %>%
  pairwise_t_test(kcals~diet, paired = TRUE,
                  p.adjust.method = "bonferroni")
pairwiseKcals
```

```{r contrast kcal 1}
#kcals.em1 <- emmeans(kcals.aov$aov, ~diet * day)

#contrast_names <- list(c(2, -1, -1, -1, -1, 2,
 #                        2, -1, -1, -1, -1, 2,
  #                       2, -1, -1, -1, -1, 2))
#contrast(kcals.em1,
 #        method = contrast_names)
```

So again, a similar story. So far, things are looking how we expected!

## Part 2

### Data Prep

Now to analyze the data from part 2! Let's read it in the same way as for the last experiment

```{r data prep 2}

column_names_p2 <- c("rat", "zero", "twenty", "forty", "sixty", "eighty", "hundred", "h2o", "condition")

datap2_grams_wide <- read_excel("IC Test Data.xlsx", sheet = 15, range = "A3:I26", col_names = column_names_p2)
datap2_kcal_wide <- read_excel("IC Test Data.xlsx", sheet = 15, range = "K3:S26", col_names = column_names_p2)

#Convert negative values to zeroes
datap2_grams_filtered <- datap2_grams_wide
datap2_grams_filtered[datap2_grams_filtered < 0] <- 0

datap2_kcal_filtered <- datap2_kcal_wide
datap2_kcal_filtered[datap2_kcal_filtered < 0] <- 0

#Drop water for now
datap2_grams_filtered2 <- datap2_grams_filtered %>%
  subset(select = -c(h2o))

datap2_kcal_filtered2 <- datap2_kcal_filtered %>%
  subset(select = -c(h2o))

#Pivot to long format
datap2_grams <- datap2_grams_filtered2 %>%
  pivot_longer(cols = diet_names, names_to = "diet", values_to = "grams")
datap2_grams$condition <- factor(datap2_grams$condition,
                           levels = c("NAL", "VEH"))
datap2_grams$diet <- factor(datap2_grams$diet,
                            levels = diet_names)
datap2_grams$rat <- factor(datap2_grams$rat,
                            levels = rat_names)

datap2_kcals <- datap2_kcal_filtered2 %>%
   pivot_longer(cols = diet_names, names_to = "diet", values_to = "kcals")
datap2_kcals$condition <- factor(datap2_kcals$condition,
                           levels = c("NAL", "VEH"))
datap2_kcals$diet <- factor(datap2_kcals$diet,
                            levels = diet_names)
datap2_kcals$rat <- factor(datap2_kcals$rat,
                            levels = rat_names)
```

### Initial Vizualizations

```{r initial viz p2}
ggboxplot(datap2_grams,
          x = "diet", y = "grams", color = "condition")

ggboxplot(datap2_kcals,
          x = "diet", y = "kcals", color = "condition")
```
Once again, the difference here is pretty stark. At least descriptively, naltrexone definitely decreased consumption, and it also functioned to sharply decrease the variability in the amount of each diet rats consumed.

### Inferential Analysis

Now for the part we've all been waiting for, the inferential analysis:

```{r hypothesis test 2 grams}

# Method 1
exp2Grams2 <- aov(grams ~ diet*condition + Error(rat/(diet*condition)), data = datap2_grams)
summary(exp2Grams2) 

# Method 2
grams.aov2 <- ezANOVA(data = datap2_grams,
                    dv = .(grams),
                    wid = .(rat),
                    within = .(diet, condition),
                    return_aov = TRUE,
                    type = 2)
grams.aov2$ANOVA
```


So we have main effects of diet and condition, but no interaction. So they consumed different amounts of the diets, and they consumed generally more on the saline day than the naltrexone day, but there wasn't a difference in the effect of naltrexone between diets. A bit of a bummer because I was hoping to dig into that interaction further, but that's real data for you.

In any case, let's do our pairwise analyses

```{r pairwise tests p2}
pairwiseGrams2 <- datap2_grams %>%
  pairwise_t_test(grams~diet, paired = TRUE,
                  p.adjust.method = "bonferroni")
pairwiseGrams2
```

```{r contrast grams 2}

#grams.em2 <- emmeans(grams.aov2$aov, ~diet * condition)

#contrast_names <- list(c(2, -1, -1, -1, -1, 2,
 #                        2, -1, -1, -1, -1, 2))
#contrast(grams.em2,
 #        method = contrast_names)
     
```

So a similar story as from part 1. I would argue in this part of the experiment the rats showed a slightly more pronounced preference for the 20% diet over the 80%, but inferentially the picture is pretty similar.

Again for completeness, let's repeat things with the calorie data:

```{r hypothesis test kcals p2}
#trying it a different way
exp2Kcals2 <- aov(kcals ~ diet*condition + Error(rat/(diet*condition)), data = datap2_kcals)
summary(exp2Kcals2) 

#And a third way because the first two don't match
kcals.aov2 <- ezANOVA(data = datap2_kcals,
                    dv = .(kcals),
                    wid = .(rat),
                    within = .(diet, condition),
                    return_aov = TRUE,
                    type = 2)
kcals.aov2$ANOVA
```

```{r pairwise tests 2 p2}
pairwiseKcals2 <- datap2_kcals %>%
  pairwise_t_test(kcals~diet, paired = TRUE,
                  p.adjust.method = "bonferroni")
pairwiseKcals2
```

```{r contrast kcal 2}
#kcals.em2 <- emmeans(kcals.aov2$aov, ~diet * condition)

#contrast_names <- list(c(2, -1, -1, -1, -1, 2,
 #                        2, -1, -1, -1, -1, 2))
#contrast(kcals.em2,
 #        method = contrast_names)
```

A similar picture as with the weight data, but here the diet x condition interaction is very much approaching significance (p = .066). Maybe something to circle back to; perhaps baseline preferences did have a small impact on the effect of naltrexone.

## Part 3

### Data Prep

```{r data prep 3}

column_names_p3 <- c("rat", "fav_blend", "fat", "sugar", "fav_blend_veh", "fav_blend_nal")
diet_names_3 <- c("fat", "sugar", "fav_blend_veh", "fav_blend_nal")

datap3_grams_wide <- read_excel("IC Test Data.xlsx", sheet = 16, range = "A3:F14", col_names = column_names_p3)
datap3_kcal_wide <- read_excel("IC Test Data.xlsx", sheet = 16, range = "I3:N14", col_names = column_names_p3)

#Convert negative values to zeroes
datap3_grams_filtered <- datap3_grams_wide
datap3_grams_filtered[datap3_grams_filtered < 0] <- 0

datap3_kcal_filtered <- datap3_kcal_wide
datap3_kcal_filtered[datap3_kcal_filtered < 0] <- 0


#Pivot to long format
datap3_grams <- datap3_grams_filtered %>%
  pivot_longer(cols = diet_names_3, names_to = "condition", values_to = "grams")
datap3_grams$condition <- factor(datap3_grams$condition,
                           levels = diet_names_3)
datap3_grams$rat <- factor(datap3_grams$rat,
                            levels = rat_names)
datap3_grams$fav_blend <- factor(datap3_grams$fav_blend,
                            levels = c("Twenty", "Eighty"))

datap3_kcals <- datap3_kcal_filtered %>%
  pivot_longer(cols = diet_names_3, names_to = "condition", values_to = "kcals")
datap3_grams$condition <- factor(datap3_grams$condition,
                           levels = diet_names_3)
datap3_grams$rat <- factor(datap3_grams$rat,
                            levels = rat_names)
datap3_grams$fav_blend <- factor(datap3_grams$fav_blend,
                            levels = c("Twenty", "Eighty"))
```

### Initial visualizations

```{r initial viz p3}
ggboxplot(datap3_grams,
          x = "condition", y = "grams", color = "condition")

ggboxplot(datap3_kcals,
          x = "condition", y = "kcals", color = "condition")
```
Things actually don't look as different between the kcal and weight data as I expected. However, I think it's going to be particularly important to present both pictures here so that the reader can draw their own conclusions.

### Inferential analyses

#### Grams

```{r hypothesis test 3 grams}

# Method 1
exp2Grams3 <- aov(grams ~ condition + Error(rat/condition), data = datap3_grams)
summary(exp2Grams3) 

# Method 2
grams.aov3 <- ezANOVA(data = datap3_grams,
                    dv = .(grams),
                    wid = .(rat),
                    within = .(condition),
                    return_aov = TRUE,
                    type = 2)
grams.aov3$ANOVA
```

And we have a significant effect of condition. Let's proceed to pairwise testing:

```{r pairwise tests grams p3}
pairwiseGrams3 <- datap3_grams %>%
  pairwise_t_test(grams~condition, paired = TRUE,
                  p.adjust.method = "bonferroni")
pairwiseGrams3
```
So according to weight, we have significant difference in all comparisons except between pure sugar and the favorite blend after naltrexone. Looks like naltrexone partially did what we expected; sugar was the less preferred pure macro, and a shot of naltrexone decreased intake of the favorite blend (the highest consumption condition here) to on par with it.

Even more surprisingly here, these rats strongly preferred fat over sugar. I *so* wish we'd given these guys DAMGO.

#### Kcals

```{r hypothesis test 3 kcals}

# Method 1
exp2Kcals3 <- aov(kcals ~ condition + Error(rat/condition), data = datap3_kcals)
summary(exp2Kcals3) 

# Method 2
kcals.aov3 <- ezANOVA(data = datap3_kcals,
                    dv = .(kcals),
                    wid = .(rat),
                    within = .(condition),
                    return_aov = TRUE,
                    type = 2)
kcals.aov3$ANOVA
```

Again, highly significant effect of condition:

```{r pairwise tests kcals p3}
pairwiseKcals3 <- datap3_kcals %>%
  pairwise_t_test(kcals~condition, paired = TRUE,
                  p.adjust.method = "bonferroni")
pairwiseKcals3
```
And here we do have a significant difference between naltrexone intake and sugar. I think this is an interesting result, but it's going to be difficult to present it. Also interestingly, there actually isn't a difference between fat intake and vehicle intake here, although I'm skeptical of this result because the fat is nearly double as caloric as the 20% diet that nearly half the rats preferred. 

I think this does it for now! I'll probably want to come back and make a few plots, but for now I know the story I'm telling!
